Implementation of Parallel Algorithms for ARIMA in Distributed Database Systems

### 2013-09-30 draft.tex
refined intro

% Introduce to MADlib, and implement ARIMA in MADLib
MADlib~\cite{madlib} is an open source library for scalable in-database
analytics.  It is created by the Predictive Analytics Team at Pivotal Inc.
MADlib provides data-parallel implementations of mathematical, statistical and
machine-learning methods for structured and unstructured data.

\section{What is MADlib?}

@@

The likelihood function $L$ for $N$ values of $Z_t$  is then
\begin{equation}
L(\phi, \theta) = \prod_{t = 1}^N  \frac{1}{\sqrt{2 \pi \sigma^2}} e^{-Z^2_t/2 \sigma^2}
\end{equation}
so the log likelihood function $l$ is
\begin{eqnarray}
l(\phi, \theta) &=& \sum_{t = 1}^N \ln \left(\frac{1}{\sqrt{2 \pi \sigma^2}}
 e^{-Z^2_t/2 \sigma^2}
 \right) \nonumber\\
 &=&  \sum_{t = 1}^N  - \ln \left( \sqrt{2 \pi \sigma^2}\right)
-\frac{Z^2_t}{2 \sigma^2}\nonumber\\
&=&  -\frac{N}{2} \ln \left( 2 \pi \sigma^2\right)  - \frac{1}{2
  \sigma^2} \sum_{t = 1}^N   Z^2_t\ \. \label{eq:loglikelihood}
\end{eqnarray}
Thus, finding the maximum likelihood is equivalent to solving the
optimization problem (known as the conditional least squares
formation)
\begin{equation}
\min_{\theta, \phi} \sum_{t = 1}^N  Z^2_t.
\end{equation}
The error term $Z_t$ can be computed iteratively as follows:
\begin{equation}
    Z_t = X_t - F_t(\phi, \theta, \mu) \label{eq:error-terms}
\end{equation}
where
\begin{equation}
F_t(\phi, \theta, \mu) = \mu + \sum_{i=1}^p \phi_i (X_{t-i}-\mu) + \sum_{i=1}^q
\theta_i Z_{t-i}
\end{equation}

>>>>

% Introduce to MADlib, and implement ARIMA in MADlib
MADlib~\cite{madlib} is an open source library for scalable in-database analytics, currently backed by the Predictive Analytics Team at Pivotal Inc.
It provides data-parallel implementations of mathematical, statistical and machine-learning methods for structured and unstructured data.

Time series analyses are important in econometrics, mathematical finance, weather forecasting, earthquake prediction and many other fields.
For example, one of the most conspicuous data analytics task -- stock price forecasting -- falls into the category of time series analysis.

However, different from other common data analytics tasks, time series data have a natural temporal ordering.
And many time series modeling methods, such as ARIMA~\cite{arima} and Cox proportional hazards~\cite{cox} therefore depend on sequential processing of time series data, which raises a challenge for the data-parallel implementation in MADlib.

In MADlib, we attack this challenge using an idea illustrated below [TODO: add a simple figure to compare MADlib ARIMA and order aggregate with iterations].

\subsection{What is MADlib?}

@@

The likelihood function $L$ for $N$ values of $Z_t$  is then
\begin{equation}
L(\phi, \theta) = \prod_{t = 1}^N  \frac{1}{\sqrt{2 \pi \sigma^2}} e^{-Z^2_t/2 \sigma^2}
\end{equation}
so the log likelihood function $l$ is
\begin{eqnarray}
l(\phi, \theta) &=& \sum_{t = 1}^N \ln \left(\frac{1}{\sqrt{2 \pi \sigma^2}}
 e^{-Z^2_t/2 \sigma^2}
 \right) \nonumber\\
 &=&  \sum_{t = 1}^N  - \ln \left( \sqrt{2 \pi \sigma^2}\right)
-\frac{Z^2_t}{2 \sigma^2}\nonumber\\
&=&  -\frac{N}{2} \ln \left( 2 \pi \sigma^2\right)  - \frac{1}{2
  \sigma^2} \sum_{t = 1}^N   Z^2_t\ . \label{eq:loglikelihood}
\end{eqnarray}
Thus, finding the maximum likelihood is equivalent to solving the
optimization problem (known as the conditional least squares
formation)
\begin{equation}
\min_{\theta, \phi} \sum_{t = 1}^N  Z^2_t.
\end{equation}
The error term $Z_t$ can be computed iteratively as follows:
\begin{equation}
    Z_t = X_t - F_t(\phi, \theta, \mu) \label{eq:error-terms}
\end{equation}
where
\begin{equation}
F_t(\phi, \theta, \mu) = \mu + \sum_{i=1}^p \phi_i (X_{t-i}-\mu) + \sum_{i=1}^q
\theta_i Z_{t-i}
\end{equation}
